\section{Software: Arcane Framework}
\label{sec:WP7:Arcane Framework:software}

\begin{itemize}
    \item \textbf{Contact Email(s):} lydie.grospellier@cea.fr
    \item \textbf{Supported Architecture(s):} HYBRID
    \item \textbf{Repository Link:} \href{https://github.com/arcaneframework/framework}{https://github.com/arcaneframework/framework}
\end{itemize}

\subsection{Software Overview}
\label{sec:WP7:Arcane Framework:summary}

Provide a brief overview of the software with respect to WP7.

\subsection{Parallel Capabilities}
\label{sec:WP7:Arcane Framework:performances}


\begin{itemize}
    \item describe the parallel programming  environment : MPI, OpenMP, CUDA, OpenACC, etc.
    \item describe the parallel computation environment: type of architecture and super computer used.
    \item describe the parallel capabilities of the software
    \item \textbf{Scalability:} Describe the general scalability properties of the software
    \item \textbf{Integration with Other Systems:} Describe how the software integrates with other numerical libraries and middleware in the Exa-MA framework.
\end{itemize}

\subsection{Initial Performance Metrics}
\label{sec:WP7:Arcane Framework:metrics}

In this section, provide a summary of the initial performance metrics of the software with respect to WP7.
You can list one or more benchmarks and their associated results, the challenges , bottlenecks and the expectations for future versions of the software.



\subsubsection{Benchmark \#1}
\begin{itemize}
    \item describe the benchmark
    \item \textbf{Benchmarking Tools Used:} Describe the tools used for parallel benchmarking and the metrics mesured
    \item \textbf{Results Summary:} Results summary not available.
    \item \textbf{Challenges Identified:} No challenges identified.
\end{itemize}

\subsection{12 months roadmap}
\label{sec:WP7:Arcane Framework:roadmap}

Describe the planned improvements  for the software in the context of WP7 and Exa-MA for the year to come that will be adressed in the next version of this deliverable.